---
title: "Gapminder Analyses Demo"
author: "Nick Kappler"
date: "10/4/2015"
output: pdf_document
---
#'Capstone' project
##Scatterplots

```{r echo=FALSE}
library(dplyr)
gap.in <- read.table("output/combined_gapMinder.tsv",
                     sep="\t",
                     header = TRUE)

gap.scand <- filter(gap.in, country == 'Finland'|country=='Sweden'|country=='Norway'|country=='Denmark')

gap.fin <- filter(gap.scand,country=='Finland')
gap.swe <- filter(gap.scand,country=='Sweden')
gap.nor <- filter(gap.scand,country=='Norway')
gap.den <- filter(gap.scand,country=='Denmark')

#Function to draw the scatter plot of an arbitrary country from gap.in
sc.plot <- function(data,state,xfactor,yfactor){
  data.country = filter(data,country==state)
  plot(x=data.country[,xfactor],y=data.country[,yfactor],
       main = c('pcGDP Time Series for ',state),
       xlab = 'year',
       ylab = 'pcGDP'
       )
}
```
Below are the per-capita gdp for Scandinavian countries.
```{r  echo=FALSE}
sc.plot(gap.in,'Finland','year','gdpPercap')
```
THis is the scatter plot for Finland.  Generally increasing, except for 1988-1990.
```{r  echo=FALSE}
sc.plot(gap.in,'Sweden','year','gdpPercap')
```
This is the scatter plot for Sweden.  Again, we have a positive trend.  The dip from 1988-1990 is less extreme than for Finland.
```{r necho=FALSE}
sc.plot(gap.in,'Norway','year','gdpPercap')
```
This ist he scatter plot for Norway.  Generally increasing, and the scale is quite a ib thigher than for the other scandinavian countries.
```{r  echo=FALSE}
sc.plot(gap.in,'Denmark','year','gdpPercap')
```
Scatter plot for Denmark. Again, note the slowdown 1988-1990 and the lower time scale than Norway.

## Summaries
```{r, data, echo=FALSE}
gap.in %>%
  select(lifeExp,continent,country)
library(dplyr)
gap.in %>% 
  select(continent,lifeExp) %>%
  group_by(continent) %>%
  summarise(mean=mean(lifeExp),min=min(lifeExp),max=max(lifeExp))
```

The mins are all very low; Oceania has a high min, but there are only two countries in the data set.  Asia has the highest maximum value, but the second lowest mean value; one of the mins is extremely low.  Oceania has teh highest average, though isn't the best category because it is only two countries.  The next highest is europe which makes sense since it is so developed.

##Histograms

hist(select(gap.in,lifeExp)[,1])
hist(select(gap.in,lifeExp)[,1],breaks=15)  
hist(select(gap.in,lifeExp)[,1],breaks=5)  